Information Aggregation in Exponential Family Markets

### 2014-04-21 WriteUp/maxentpredmkts.tex
took a pass and made small edits and typo fixes

\textbf{Related Work.}
The notion of an exponential family distribution is fundamental to this paper. For comprehensive introductions to these distributions, see~\citep{Barndorff78,WainJordan08}. Exponential families are intimately tied to the notions of log loss and entropy, but can be generalized to other types of convex losses and information, as shown by~\citet{GrunwaldDawid}, who also make a connection to scoring rules.

@@

Prediction markets are known to perform well in practice~\citep{AGT07,pennock2001real}. However, a sound theory for interpreting trader behavior and market prices is an ongoing area of study~\citep{wolfers2006interpreting}. At one extreme, agents are assumed myopic and risk-neutral, implying they move the market state to their belief~\citep{chen2010new}. At the other extreme, agents are strategic and the market fully incorporates all information~\citep{ostrovsky2012}. 
% Strategic behavior in prediction markets is addressed in~\citep{chen2010gaming,dimitrov08}.

We are not aware of any works that consider risk-averse agents within cost-function based markets. However, risk aversion is a fundamental component of mathematical finance and portfolio optimization, and there are close connections between the notion of a cost function and that of a convex risk measure~\citep{follmerRisk,follmer2011entropic}. Indeed, they arise from the same axioms as noted by~\citet{othman2011liqu}. We see the potential to draw more on the mathematical finance literature to take into account risk aversion, as prediction markets are simply single-period financial markets~\citep[Part I]{follmerFinance}. We also note that connections between Machine Learning and market mechanisms have been explored in \cite{storkey11}.

@@

When the outcome space is finite we recover classical scoring rules by using the statistic $\phi : \mX \rightarrow \{0,1\}^{\mX}$ that maps an outcome $x$ to a unit vector with a 1 in the component corresponding to $x$. The expectation of $\phi$ is then exactly the probability mass function.

@@

It is well-known that solutions to~(\ref{maxent-prog}) are \emph{exponential family} distributions (see, for instance, \cite{WainJordan08}). These are distributions whose densities with respect to $\nu$ take the form
%
\begin{equation} \label{exp-fam}
p(x;\theta) = \exp( \la \theta, \phi(x) \ra - T(\theta) ).
\end{equation}
%
The density is stated here in terms of its \emph{natural} parametrization $\theta \in \bR^d$, where $\theta$ arises as the Lagrange multiplier associated with the linear constraints in~(\ref{maxent-prog}). The term $T(\theta)$ essentially arises as the multiplier for the normalization constraint (the density must integrate to 1), and so ensures that~(\ref{exp-fam}) is normalized:
%
\begin{equation} \label{log-part}
T(\theta) = \log \int_{\mX} \exp \la \theta, \phi(x) \ra \,d\nu(x).
\end{equation}
%
The function $T$ is known as the \emph{log-partition} or \emph{cumulant} function corresponding to the exponential family. Its domain is $\Theta = \{ \theta \in \bR^d : T(\theta) < +\infty \}$, called the natural parameter space, and we assume throughout that it is nonempty.\footnote{This may impose additional conditions on the choice of base measure $\nu$ in certain special cases. For instance, if the density over $\bR$ is known to have unit variance and statistic $\phi(x) = x$, then one uses the base measure $\nu(x) = e^{-x^2/2}$ leading to a normal distribution with fixed variance~\citep{Barndorff78}.}
%
The exponential family is \emph{regular} if $\Theta$ is open---almost all exponential families of interest, and all those we consider in this work, are regular. The family is \emph{minimal} if there is no $\alpha \in \Theta$ such that $\la \alpha, \phi(x) \ra$ is a constant over $\mX$ ($\nu$-almost everywhere); minimality is a property of the associated statistic $\phi$, usually called the \emph{sufficient statistic} in the literature.

Some commonly studied distributions that are in fact exponential families include the binomial, beta and normal distributions. To see for instance, that the normal distribution is an exponential family note that instead of the usual parametrization using mean  $\mu$ and  variance $\sigma$, it can be parametrized in terms of $\vec{\theta}=(\theta_{1},\theta_{2})$ where $\theta_{1}=\frac{\mu}{\sigma^{2}}$ and $\theta_{2}=\frac{-1}{2\sigma^{2}}$,  with corresponding log partition function $-\frac{\beta_{1}^{2}}{4\beta_{2}}-\frac{1}{2}\log(-2\beta_{2})$.

The following proposition collects the relevant results on regular exponential families; proofs may be found in~\citet[Prop.\ 3.1--3.2, Thm.\ 3.3--3.4]{WainJordan08} and see also~\citet[Lem.\ 1, Thm.\ 2]{BanerjeeDhGh05b}. A convex function $T$ is of \emph{Legendre type} if it is proper, closed, strictly convex and differentiable on the interior of its domain, and $\lim_{\theta \rightarrow \bar{\theta}} \norm{\grad T(\theta)} = +\infty$ when $\bar{\theta}$ lies on the boundary of the domain. 
%
\begin{prop}\label{prop:exp}
Consider a regular exponential family with minimal sufficient statistic. The following properties hold:
\begin{enumerate}
\item $T$ and $G$ are of Legendre type, and $T = G^*$ (equivalently $G = T^*$). 
\item The gradient map $\grad T$ is one-to-one and onto the interior of $\mM$. Its inverse is $\grad G$ which is one-to-one and onto the interior of $\Theta$.
\item The exponential family distribution with natural parameter $\theta \in \Theta$ has expected statistic $\mu = \Exp_p[\phi(x)] = \grad T(\theta)$.
\item The maximum entropy distribution with expected statistic $\mu$ is the exponential family distribution with natural parameter $\theta = \grad G(\mu)$. 
\end{enumerate}
\end{prop}
%
In the above $T^*$ denotes the convex conjugate of $T$, which here can be evaluated as $T^*(\mu) = \sup_{\theta \in \Theta} \la \theta, \mu \ra - T(\theta)$. Similarly, $G^*(\theta) = \sup_{\mu \in \mM} \la \theta, \mu \ra - G(\mu)$.

@@

The following observation is crucial. Let $\tp \in \mP$ be a density (not necessarily from an exponential family) with expected statistic $\mu$, let $p(\cdot\,;\mu)$ be the exponential family with the same expected statistic, and let $\hmu \in \mM$ be an alternative report. Then from~(\ref{exp-fam}) note
%
\begin{equation} \label{equalizer-rule}
\Exp_{\tp}[\log p(x;\hmu)] = \Exp_{p(\cdot;\mu)}[\log p(x;\hmu)] = \la \htheta, \mu \ra - T(\htheta),
\end{equation}
%
where $\htheta = \grad G(\hmu)$ is the natural parameter for the exponential family with statistic $\hmu$. Note that the expected log score does not depend  on the entire maximum entropy distribution but just its expected statistic $\mu$. This is how we can achieve proper scoring according to Definition~{\ref{def:proper-score}.
%We see from this that the expected log score only depends on the expectation $\mu$ of the underlying density, not the full density, which is how we can achieve proper scoring according to Definition~{\ref{def:proper-score}.
%
\begin{theorem} \label{maxent-score}
Consider the logarithmic scoring rule $S(\mu,x) = \log p(x;\mu)$ defined over a set of densities $\mD$ parametrized by $\mM$. The scoring rule is proper if and only if $\mD$ is the exponential family with statistic $\phi$, where $\mu=\E_{p(x;\mu)}[\phi(x)]$. 
\end{theorem}
%
%
\begin{proof}
Let $\mu,\hmu \in \mM$ be the agent's true belief and an alternative report, and let $p \in \mP$ be a density consistent with $\mu$. Let $\theta = \grad G(\mu)$ and $\htheta = \grad G(\hmu)$, and note that $\mu = \grad T(\theta)$. We have
%
\begin{eqnarray}
\Exp_p[\log p(x;\mu)] - \Exp_p[\log p(x;\hmu)]
& = & \la \theta, \mu \ra - T(\theta) - \la \htheta, \mu \ra + T(\htheta) \nonumber\\
& = & T(\htheta) - T(\theta) - \la \htheta - \theta, \mu \ra \nonumber\\
& = & T(\htheta) - T(\theta) - \la \htheta - \theta, \grad T(\theta) \ra. \label{breg-div}
\end{eqnarray}
%
The latter is positive by the strict convexity of $T$, which shows that the log score is proper. For the converse, assume the defined log score is proper. By the Savage characterization of proper scoring rules for expectations (see~\citet[Thm.\ 11]{AbernethyFr12}), we must have
%
$$
S(\mu,x) = G(\mu) - \la \grad G(\mu), \mu - \phi(x) \ra
$$
for some strictly convex function $G$. Let $T = G^*$, so that $\grad G = {\grad T}^{-1}$, and let $\theta = \grad G(\mu)$. Then the above can be written as
%
\begin{eqnarray*}
\log p(x;\mu) & = & G(\mu) - \la \grad G(\mu), \mu - \phi(x) \ra \\
& = & \la \theta, \mu \ra - T(\theta) - \la \theta, \mu - \phi(x) \ra = \la \theta, \phi(x) \ra - T(\theta),
\end{eqnarray*}
%
which shows that $p(x;\mu)$ takes the form of an exponential family.
\end{proof}
%
It has been understood since~\citep{mccarthy1956measures,Savage71} that scoring rules for densities and expectations can be characterized in terms of an underlying convex function. The generalized log scoring rule of Theorem~\ref{maxent-score} obtains by applying the modern characterization of~\citet{AbernethyFr12} to the optimal value function $G$ (negative maximum entropy). \citet[Thm.\ 6]{BanerjeeDhGh05} prove a similar result by showing a bijection between exponential families and Bregman divergences.

@@

\subsection{Examples: Moments over the Real Line}

Theorem~\ref{maxent-score} leads to a straightforward procedure for constructing score rules for expectations. Define the relevant statistic, and consider the maximum entropy (equivalently, exponential family) distribution consistent with the agent's reported mean $\mu$. The scoring rule compensates the agent according to the log likelihood of the eventual outcome according to this distribution. The interpretation is that the agent is only providing partial information about the underlying density, so the principal first infers a full density according to the principle of maximum entropy, and then scores the agent using the standard log score. We stress that the exponential families framework is a tool in the construction of scoring rules; we are not actually modeling any agent's belief via an exponential family.

An advantage of this generalization of the log score is that, for many domains (multi-dimensional included) and expectations of interest, it leads to novel closed-form scoring rules. By examining the log densities of various exponential families, we can for instance obtain scoring rules for several different combinations of the arithmetic, geometric, and harmonic means, as well as higher order moments. The following examples illustrate the construction. Here, we have used $\mu$ and $\sigma$ for the true mean and variance and $\hmu$ and $\hsigma$ for the elicited mean and variance respectively (and $\hSigma$ for the variance of the multivariate normal).

%
\begin{example}
As a base measure we take the counting measure over the finite set $\{1,2,\ldots,d\}$, and we consider the statistic $\phi(i) \in \{0,1\}^d$ that maps outcome $i$ to the unit vector with component $i$ equal to 1. The expectation of $\phi(i)$ is simply the entire probability distribution $p$, which is a multinomial distribution (an instance of an exponential family). We recover the standard log scoring rule $S(\hat{p}, i) = \log \hat{p}(i)$. 
\end{example}
%

@@

As a concrete example, recall that in the classic finite-outcome case the statistic has a component for each outcome $x$ such that $\phi_x(x') = 1$ if $x' = x$ and 0 otherwise. Therefore the corresponding security pays 1 dollar if outcome $x$ occurs. (These are known as Arrow-Debreu securities.) In Example~\ref{ex:exponential} the one-dimensional statistic is $\phi(x) = x$, corresponding to a security whose payoff is linear in the outcome $x \in \bR_+$. (This amounts to a futures contract.)

@@

There is a well-known duality between scoring rules and cost-function based markets~\citep{acv13,H03}. To see this in our context, recall from~(\ref{equalizer-rule}):
%
\begin{equation*}
\Exp_{\tp}[\log p(x;\hmu)] = \la \htheta, \mu \ra - T(\htheta)
\end{equation*}
%
where $\tp$ is the agent's belief and $\hmu$ the agent's report. The expected log score from reporting $\hmu$ is exactly the same as the expected payoff from buying portfolio of shares $\htheta = \grad C(\hmu)$ (assuming an initial market state of 0), as $\la \htheta, \mu \ra$ is the expected revenue and $T(\htheta)$ is the cost. As in Section~\ref{sec:scoring} this reasoning relies on the assumption of risk-neutrality, not on any specific form for the agent's belief.

@@

Let us now revisit our scoring rule examples from Section~\ref{sec:scoring} in the context of prediction markets. The relevant entities now are the payoff function, the effective domain of shares, and the cost function.
%
\begin{example}
We consider a finite outcome space $\mX = \{1,2,\ldots,d\}$ and set up a market with an Arrow-Debreu security $\phi_i$ for every outcome $i$: the security pays off 1 if $i$ occurs and 0 otherwise. The log-partition function recovers the classical LMSR cost function~\citep{H07}:
$C(\theta) = \log \sum_{i \in \mX} e^{\theta_i}.$
The effective domain is $\Theta = \bR^d$, meaning that a security can always be bought or sold short.
\end{example}
%

@@

The exponential families framework is well-suited to reasoning about Bayesian updates. As before let the data distribution be given by $p(x;\theta)=\exp(\la\theta,\phi(x)\ra-T(\theta))$ where $T$ is the log partition function and $\phi$ are the sufficient statistics. Instead of direct beliefs about the data distribution the agent maintains a conjugate prior over the parameters $\theta$. Every exponential family admits a conjugate prior of the form
%
$$p(\theta;b_{0})=\exp(\la n\nu,\theta\ra + nT(\theta) - \psi(\nu,n)).$$
%
Note that this is also an exponential family with natural parameter $b_0 = (n\nu,n)$ where $\nu \in \bR^d$ and $n$ is a positive integer. The sufficient statistic maps $\theta$ to $(\theta, T(\theta))$, and the log partition function $\psi$ is defined as the normalizer as usual. For a complete treatment of exponential families conjugate priors, see for instance~\citet{Barndorff78}. Now~\citet[Thm.\ 2]{diaconis79} and~\citet{jewell1974credible} have shown that
%
\begin{equation}\label{eq:prior}
\E_{\theta\sim b_{0}}\E_{x\sim \theta}[\phi(x)]=\nu,
\end{equation}
%
meaning that $\nu = n \nu / n$ is the posterior mean. Thus, it is helpful to think of the prior as being based on a `phantom' sample of size $n$ and mean $\nu$. Suppose now that the agent observes an empirical sample with mean $\hmu$ and size $m$. By a standard derivation~\citep[see][]{diaconis79}, the posterior conjugate prior parameters become $n\nu \leftarrow n\nu + m\hmu$ and $n \leftarrow n+m$, and the posterior expectation~(\ref{eq:prior}) evaluates to
%
\begin{equation} \label{posterior}
\frac{n\nu + m\hmu}{n+m}.
\end{equation}
%
Thus the posterior mean is a convex combination of the prior and empirical means, and their relative weights depend on the phantom and empirical sample sizes.% need a reference here

@@

\noindent
Recall that for the generalized LMSR, the cost function $C$ is exactly the log partition function $T$. We are therefore lead to the following understanding of a risk-averse agent's behavior in such a market.
%
\begin{theorem} \label{thm:exp-util}
Suppose an agent has exponential utility with coefficient $a$ and exponential family beliefs with natural parameter $\htheta$. In the generalized LMSR market with current market state $\theta$, the agent's optimal trade $\delta$ moves the state vector to
%
\begin{equation} \label{exp-update}
\theta + \delta = \frac{1}{1+a} \htheta + \frac{a}{1+a} \theta.
\end{equation}
%
\end{theorem}
%
%
\begin{proof}
The agent's optimal trade maximizes its expected utility, or equivalently the certainty equivalent. From Lemma~\ref{lem:exp-util} and that $T = C$, the agent maximizes
%
$$
\log a - T(\htheta - a\delta) + T(\htheta) - aT(\theta+\delta) + aT(\theta). 
$$
This objective is strictly concave, from the strict convexity of $T$. The optimum is therefore characterized by the first-order condition $\grad T(\htheta - a\delta) = \grad T(\theta+\delta)$. As the gradient map $\grad T$ is one-to-one, this is solved by equating the arguments, which leads to $\delta = (\htheta - \theta) / (1+a)$ and~(\ref{exp-update}).
% 
\end{proof}
%
We stress that the aggregation in Theorem~\ref{thm:exp-util} arises directly from the agents' choice of trades. The market maker does not need to know the risk aversion parameters. Note that as, $a$ tends to 0, we approach risk neutrality and the agent moves the share vector all the way to its private estimate $\htheta$. As $a$ grows larger (the agent grows more risk averse) the agent makes smaller trades to reduce it exposure, and the final state stays closer to the current state $\theta$. Update~(\ref{exp-update}) implies that, under the conditions of the theorem, a market that receives a sequence of myopic traders aggregates their natural parameters in the form of an exponentially weighted moving average. The final market estimates (i.e., prices) are obtained by applying $\grad T$ to this average.

@@

In this section, we consider the evolution of the market state when traders are budget-limited.
% and their budgets are known to the market maker. We consider a cost-function based market with cost function $C$ and payoff $\phi(x)$ with non-negative payoff. 
We assume that the traders trade in multiple instances of the market. That is, after the initial setup and trades, the outcome is revealed and the traders are paid off. Then, the market runs again with the same set of traders and the outcome drawn from a (possibly different) exponential family distribution. The goal of this section is to analyze how the budget of traders evolve under these circumstances. As before, the market price is interpreted as a probability density over the outcome space and the share vector as the natural parameter of an exponential family distribution. Consistent with the connections drawn in Section~\ref{sec:scoring} and throughout, we measure the error in prediction using the standard log loss.

@@

Note that while an uninformative trader is unconstrained in his belief, the informative trader's belief corresponds to the parameter of the true distribution.
\paragraph{Limited Damage}
We will now quantify the error in prediction that the market maker might have to endure as a result of ill-informed entities entering the market. We assume that these entities trade in multiple instances of the market; thus the exposure of the market maker is over several rounds. The log loss function for $\theta$ shares held is defined as $L(\theta,x)=-\log p(x;\theta)=C(\theta)-\la \theta\phi(x)\ra$.

>>>>

\textbf{Related Work.}
The notion of an exponential family distribution is fundamental to this paper. For comprehensive introductions to these distributions, see~\citep{Barndorff78,WainJordan08}. Exponential families are intimately tied to the notions of log loss and entropy, but can be generalized to other types of convex losses and information as shown by~\citet{GrunwaldDawid}, who also make a connection to scoring rules.

@@

Prediction markets are known to perform well in practice~\citep{AGT07,pennock2001real}. However, a sound theory for interpreting trader behavior and market prices is an ongoing area of study~\citep{wolfers2006interpreting}. At one extreme, agents are assumed myopic and risk-neutral, implying they move the market state to their belief~\citep{chen2010new}. At the other extreme, agents are strategic and the market fully incorporates all information~\citep{ostrovsky2012}. 
% Strategic behavior in prediction markets is addressed in~\citep{chen2010gaming,dimitrov08}.
The aggregation properties of market mechanisms have also been explored from a machine learning perspective in~\cite{storkey11}.

We are not aware of any works that consider risk-averse agents within cost-function based markets. However, risk aversion is a fundamental component of mathematical finance and portfolio optimization, and there are close connections between the notion of a cost function and that of a convex risk measure~\citep{follmerRisk,follmer2011entropic}. Indeed, they arise from the same axioms as noted by~\citet{othman2011liqu}. We see the potential to draw more on the mathematical finance literature to take into account risk aversion, as prediction markets are simply single-period financial markets~\citep[Part I]{follmerFinance}.

@@

%When the outcome space is finite we recover classical scoring rules by using the statistic $\phi : \mX \rightarrow \{0,1\}^{\mX}$ that maps an outcome $x$ to a unit vector with a 1 in the component corresponding to $x$. The expectation of $\phi$ is then exactly the probability mass function.

@@

It is well-known that solutions to~(\ref{maxent-prog}) are \emph{exponential family} distributions; see, for instance \cite{WainJordan08}. These are distributions whose densities with respect to $\nu$ take the form
%
\begin{equation} \label{exp-fam}
p(x;\theta) = \exp( \la \theta, \phi(x) \ra - T(\theta) ).
\end{equation}
%
The density is stated here in terms of its \emph{natural} parametrization $\theta \in \bR^d$, where $\theta$ arises as the Lagrange multiplier associated with the linear constraints in~(\ref{maxent-prog}). The term $T(\theta)$ essentially arises as the multiplier for the normalization constraint (the density must integrate to 1), and so ensures that~(\ref{exp-fam}) is normalized:
%
\begin{equation} \label{log-part}
T(\theta) = \log \int_{\mX} \exp \la \theta, \phi(x) \ra \,d\nu(x).
\end{equation}
%
The function $T$ is known as the \emph{log-partition} or \emph{cumulant} function corresponding to the exponential family. Its domain is $\Theta = \{ \theta \in \bR^d : T(\theta) < +\infty \}$, called the natural parameter space, and we assume throughout that it is nonempty.\footnote{This may impose additional conditions on the choice of base measure $\nu$ in certain special cases. For instance, if the density over $\bR$ is known to have unit variance and statistic $\phi(x) = x$, then one uses the base measure $\nu(x) = e^{-x^2/2}$ leading to a normal distribution with fixed variance~\citep{Barndorff78}.}
%
The exponential family is \emph{regular} if $\Theta$ is open---almost all exponential families of interest, and all those we consider in this work, are regular. The family is \emph{minimal} if there is no $\alpha \in \Theta$ such that $\la \alpha, \phi(x) \ra$ is a constant over $\mX$ ($\nu$-almost everywhere); minimality is a property of the associated statistic $\phi$, usually called the \emph{sufficient statistic} in the literature.

Some commonly studied distributions that are in fact exponential families include the binomial, beta, Poisson, exponential, and normal distributions; the sufficient statistics and parametrizations for the exponential and normal will be covered in Section~\ref{sec:examples}.
%To see for instance, that the normal distribution is an exponential family note that instead of the usual parametrization using mean  $\mu$ and  variance $\sigma$, it can be parametrized in terms of $\vec{\theta}=(\theta_{1},\theta_{2})$ where $\theta_{1}=\frac{\mu}{\sigma^{2}}$ and $\theta_{2}=\frac{-1}{2\sigma^{2}}$,  with corresponding log partition function $-\frac{\beta_{1}^{2}}{4\beta_{2}}-\frac{1}{2}\log(-2\beta_{2})$. [SL: I think this is too much so removed it.]
%
The following proposition collects the relevant results on regular exponential families; proofs may be found in~\citet[Prop.\ 3.1--3.2, Thm.\ 3.3--3.4]{WainJordan08} and see also~\citet[Lem.\ 1, Thm.\ 2]{BanerjeeDhGh05b}. A convex function $T$ is of \emph{Legendre type} if it is proper, closed, strictly convex and differentiable on the interior of its domain, and $\lim_{\theta \rightarrow \bar{\theta}} \norm{\grad T(\theta)} = +\infty$ when $\bar{\theta}$ lies on the boundary of the domain. 
%
\begin{prop}\label{prop:exp}
Consider a regular exponential family with minimal sufficient statistic. The following properties hold:
\begin{enumerate}
\item $T$ and $G$ are of Legendre type, and $T = G^*$ (equivalently $G = T^*$). 
\item The gradient map $\grad T$ is one-to-one and onto the interior of $\mM$. Its inverse is $\grad G$ which is one-to-one and onto the interior of $\Theta$.
\item The exponential family distribution with natural parameter $\theta \in \Theta$ has expected statistic $\mu = \Exp_p[\phi(x)] = \grad T(\theta)$.
\item The maximum entropy distribution with expected statistic $\mu$ is the exponential family distribution with natural parameter $\theta = \grad G(\mu)$. 
\end{enumerate}
\end{prop}
%
In the above $T^*$ denotes the convex conjugate of $T$, which here can be evaluated as $T^*(\mu) = \sup_{\theta \in \Theta} \la \theta, \mu \ra - T(\theta)$. Similarly, $G^*(\theta) = \sup_{\mu \in \mM} \la \theta, \mu \ra - G(\mu)$.

@@

The following observation is crucial. Let $\tp \in \mP$ be a density (not necessarily from an exponential family) with expected statistic $\mu$, let $p(\cdot\,;\mu)$ be the exponential family with the same expected statistic, and let $\hmu \in \mM$ be an alternative report. Then from~(\ref{exp-fam}) we have
%
\begin{equation} \label{equalizer-rule}
\Exp_{\tp}[\log p(x;\hmu)] = \Exp_{p(\cdot;\mu)}[\log p(x;\hmu)] = \la \htheta, \mu \ra - T(\htheta),
\end{equation}
%
where $\htheta = \grad G(\hmu)$ is the natural parameter for the exponential family with statistic $\hmu$. Note that the expected log score does not depend  on the entire maximum entropy distribution but just its expected statistic $\mu$. This is how we can achieve proper scoring according to Definition~{\ref{def:proper-score}.
%We see from this that the expected log score only depends on the expectation $\mu$ of the underlying density, not the full density, which is how we can achieve proper scoring according to Definition~{\ref{def:proper-score}.
%
\begin{theorem} \label{maxent-score}
Consider the logarithmic scoring rule $S(\mu,x) = \log p(x;\mu)$ defined over a set of densities $\mD$ parametrized by $\mM$. The scoring rule is proper if and only if $\mD$ is the exponential family with statistic $\phi$, where $\mu=\E_{p(x;\mu)}[\phi(x)]$. 
\end{theorem}
%
%
\begin{proof}
Let $\mu,\hmu \in \mM$ be the agent's true belief and an alternative report, and let $p \in \mP$ be a density consistent with $\mu$. Let $\theta = \grad G(\mu)$ and $\htheta = \grad G(\hmu)$, and note that $\mu = \grad T(\theta)$. We have
%
\begin{eqnarray}
\Exp_p[\log p(x;\mu)] - \Exp_p[\log p(x;\hmu)]
& = & \la \theta, \mu \ra - T(\theta) - \la \htheta, \mu \ra + T(\htheta) \nonumber\\
& = & T(\htheta) - T(\theta) - \la \htheta - \theta, \mu \ra \nonumber\\
& = & T(\htheta) - T(\theta) - \la \htheta - \theta, \grad T(\theta) \ra. \label{breg-div}
\end{eqnarray}
%
The latter is positive by the strict convexity of $T$, which shows that the log score is proper. For the converse, assume the defined log score is proper. By the Savage characterization of proper scoring rules for expectations (see~\citet[Thm.\ 11]{AbernethyFr12}), we must have
%
$$
S(\mu,x) = G(\mu) - \la \grad G(\mu), \mu - \phi(x) \ra
$$
for some strictly convex function $G$. Let $T = G^*$, so that $\grad G = {\grad T}^{-1}$, and let $\theta = \grad G(\mu)$. Then the above can be written as
%
\begin{eqnarray*}
\log p(x;\mu) & = & G(\mu) - \la \grad G(\mu), \mu - \phi(x) \ra \\
& = & \la \theta, \mu \ra - T(\theta) - \la \theta, \mu - \phi(x) \ra = \la \theta, \phi(x) \ra - T(\theta),
\end{eqnarray*}
%
which shows that $p(x;\mu)$ takes the form of an exponential family.
\end{proof}
%
It has been understood since~\citep{mccarthy1956measures,Savage71} that scoring rules for densities and expectations can be characterized in terms of an underlying convex function. The generalized log scoring rule of Theorem~\ref{maxent-score} obtains by applying the modern characterization of~\citet[Thm.\ 11]{AbernethyFr12} to the optimal value function $G$ (negative maximum entropy). \citet[Thm.\ 6]{BanerjeeDhGh05} prove a similar result by showing a bijection between exponential families and Bregman divergences.

@@

\subsection{Examples: Moments over the Real Line}
\label{sec:examples}

Theorem~\ref{maxent-score} leads to a straightforward procedure for constructing score rules for expectations. Define the relevant statistic, and consider the maximum entropy (equivalently, exponential family) distribution consistent with the agent's reported mean $\mu$. The scoring rule compensates the agent according to the log likelihood of the eventual outcome according to this distribution. The interpretation is that the agent is only providing partial information about the underlying density, so the principal first infers a full density according to the principle of maximum entropy, and then scores the agent using the standard log score. We stress that the exponential families framework is a tool in the construction of scoring rules at this point; we are not actually modeling any agent's belief via an exponential family.

An advantage of this generalization of the log score is that, for many domains (multi-dimensional included) and expectations of interest, it leads to novel closed-form scoring rules. By examining the log densities of various exponential families, we can for instance obtain scoring rules for several different combinations of the arithmetic, geometric, and harmonic means, as well as higher order moments. The following examples illustrate the construction. Here, we have used $\mu$ and $\sigma$ for the true mean and variance and $\hmu$ and $\hsigma$ for the elicited mean and variance respectively (and $\hSigma$ for the covariance matrix of the multivariate normal).

%
\begin{example} \label{ex:multinomial}
As a base measure we take the counting measure over the finite set $\{1,2,\ldots,d\}$, and we consider the statistic $\phi(i) \in \{0,1\}^d$ that maps outcome $i$ to the unit vector with component $i$ equal to 1. The expectation of $\phi(i)$ is simply the entire probability distribution $p$, which is a multinomial distribution (an instance of an exponential family). We recover the standard log scoring rule $S(\hat{p}, i) = \log \hat{p}(i)$. 
\end{example}
%

@@

To be concrete, recall from Example~\ref{ex:multinomial} that in the classic finite-outcome case the statistic has a component for each outcome $x$ such that $\phi_x(x') = 1$ if $x' = x$ and 0 otherwise. Therefore the corresponding security pays 1 dollar if outcome $x$ occurs. (These are known as Arrow-Debreu securities.) In Example~\ref{ex:exponential} the one-dimensional statistic is $\phi(x) = x$, corresponding to a security whose payoff is linear in the outcome $x \in \bR_+$. (This amounts to a futures contract.)

@@

There is a well-known duality between scoring rules and cost-function based markets~\citep{acv13,H03}. To see this in our context, recall from~(\ref{equalizer-rule}) that
%
\begin{equation*}
\Exp_{\tp}[\log p(x;\hmu)] = \la \htheta, \mu \ra - T(\htheta)
\end{equation*}
%
where $\tp$ is the agent's belief and $\hmu$ the agent's report. The expected log score from reporting $\hmu$ is exactly the same as the expected payoff from buying portfolio of shares $\htheta = \grad C(\hmu)$ (assuming an initial market state of 0), as $\la \htheta, \mu \ra$ is the expected revenue and $T(\htheta)$ is the cost. As in Section~\ref{sec:scoring} this reasoning relies on the assumption of risk-neutrality, not on any specific form for the agent's belief.

@@

Let us now revisit our scoring rule examples from Section~\ref{sec:scoring} in the context of prediction markets. The relevant entities now are the payoff function, the effective domain of shares, and the cost function.
%
\begin{example}
We consider a finite outcome space $\mX = \{1,2,\ldots,d\}$ and set up a market with an Arrow-Debreu security $\phi_i$ for every outcome $i$: the security pays off 1 if $i$ occurs and 0 otherwise. The log-partition function recovers the classical LMSR cost function:
$C(\theta) = \log \sum_{i \in \mX} e^{\theta_i}.$
The effective domain is $\Theta = \bR^d$, meaning that a security can always be bought or sold short.
\end{example}
%

@@

The exponential families framework is well-suited to reasoning about Bayesian updates. As before let the data distribution be given by $p(x;\theta)=\exp(\la\theta,\phi(x)\ra-T(\theta))$ where $T$ is the log partition function and $\phi$ are the sufficient statistics. Instead of direct beliefs about the data distribution, the agent maintains a conjugate prior over the parameters $\theta$. Every exponential family admits a conjugate prior of the form
%
$$p(\theta;n\nu, n)=\exp(\la n\nu,\theta\ra + nT(\theta) - \psi(\nu,n)).$$
%
Note that this is also an exponential family with natural parameter $(n\nu,n)$ where $\nu \in \bR^d$ and $n$ is a positive integer. The sufficient statistic maps $\theta$ to $(\theta, T(\theta))$, and the log partition function $\psi$ is defined as the normalizer as usual. For a complete treatment of exponential families conjugate priors, see for instance~\citep{Barndorff78}. Now~\citet[Thm.\ 2]{diaconis79} and~\citet{jewell1974credible} have shown that
%
\begin{equation}\label{eq:prior}
\E_{\theta\sim (n\nu,n)}\E_{x\sim \theta}[\phi(x)]=\nu,
\end{equation}
%
meaning that $\nu = n \nu / n$ is the posterior mean. Thus, it is helpful to think of the prior as being based on a `phantom' sample of size $n$ and mean $\nu$. Suppose now that the agent observes an empirical sample with mean $\hmu$ and size $m$. By a standard derivation~\citep[see][]{diaconis79}, the posterior conjugate prior parameters become $n\nu \leftarrow n\nu + m\hmu$ and $n \leftarrow n+m$, and the posterior expectation~(\ref{eq:prior}) evaluates to
%
\begin{equation} \label{posterior}
\frac{n\nu + m\hmu}{n+m}.
\end{equation}
%
Thus the posterior mean is a convex combination of the prior and empirical means, and their relative weights depend on the phantom and empirical sample sizes.% need a reference here

@@

\noindent
Recall that for the generalized LMSR, the cost function $C$ is exactly the log partition function $T$. We are therefore lead to the following understanding of a risk-averse agent's behavior in such a market.
%
\begin{theorem} \label{thm:exp-util}
Suppose an agent has exponential utility with coefficient $a$ and exponential family beliefs with natural parameter $\htheta$. In the generalized LMSR market with current market state $\theta$, the agent's optimal trade $\delta$ moves the state vector to
%
\begin{equation} \label{exp-update}
\theta + \delta = \frac{1}{1+a} \htheta + \frac{a}{1+a} \theta.
\end{equation}
%
\end{theorem}
%
%
\begin{proof}
The agent's optimal trade maximizes its expected utility, or equivalently the certainty equivalent. From Lemma~\ref{lem:exp-util} and that $T = C$, the agent maximizes
%
$$
\log a - T(\htheta - a\delta) + T(\htheta) - aT(\theta+\delta) + aT(\theta). 
$$
This objective is strictly concave, from the strict convexity of $T$. The optimum is therefore characterized by the first-order condition $\grad T(\htheta - a\delta) = \grad T(\theta+\delta)$. As the gradient map $\grad T$ is one-to-one, this is solved by equating the arguments, which leads to $\delta = (\htheta - \theta) / (1+a)$ and~(\ref{exp-update}).
% 
\end{proof}
%
We stress that the aggregation in Theorem~\ref{thm:exp-util} arises directly from the agents' choice of trades. The market maker does not need to know the risk aversion parameters. Note that as $a$ tends to 0 we approach risk neutrality and the agent moves the share vector all the way to its private estimate $\htheta$. As $a$ grows larger (the agent grows more risk averse) the agent makes smaller trades to reduce it exposure, and the final state stays closer to the current state $\theta$. Update~(\ref{exp-update}) implies that, under the conditions of the theorem, a market that receives a sequence of myopic traders aggregates their natural parameters in the form of an exponentially weighted moving average. The final market estimates (i.e., prices) are obtained by applying $\grad T$ to this average.

@@

In this section, we consider the evolution of the market state when traders are budget-limited.
% and their budgets are known to the market maker. We consider a cost-function based market with cost function $C$ and payoff $\phi(x)$ with non-negative payoff. 
We assume that the traders trade in multiple instances of the market. That is, after the initial setup and trades, the outcome is revealed and the traders are paid off. Then, the market runs again with the same set of traders and the outcome drawn from a (possibly different) exponential family distribution. The goal of this section is to analyze how the budgets of traders evolve under these circumstances. As before, the market price is interpreted as a probability density over the outcome space and the share vector as the natural parameter of an exponential family distribution. Consistent with the connections drawn in Section~\ref{sec:scoring} and throughout, we measure the error in prediction using the standard log loss.

@@

Note that while an uninformative trader is unconstrained in his belief, the informative trader's belief corresponds to the parameter of the true distribution.
\paragraph{Limited Damage}
We will now quantify the error in prediction that the market maker might have to endure as a result of ill-informed entities entering the market. We assume that these entities trade in multiple instances of the market; thus the exposure of the market maker is over several rounds. The log loss function for $\theta$ shares held is defined as $L(\theta,x)=-\log p(x;\theta)=C(\theta)-\la \theta,\phi(x)\ra$.
